--- 
title: "Useful Tips"
author: "Lingyun Zhang"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
link-citations: yes
links-as-notes: true
colorlinks: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preface {-}

This book contains useful Tex, R and Python tips.

# Tex tips {-}

## Commenting in WinEdt

In Winedt use `Ctrl+Shift+Alt+→` to comment a line and to uncomment a line use `Ctrl+Shift+Alt+←`.

**Source**: https://tex.stackexchange.com/questions/230997/which-windows-editors-have-comment-uncomment-keyboard-shortcuts


## Rotate a table

**Q**: How to rotate a table in latex?

**A**:
````
\usepackage{rotating}
\begin{sidewaystable}[h]
\centering
\caption{Mean, Median and Standard Deviation at AU level from Simulations, where
                  tolerance limit set at MB level and $p=0.04$}
\label{table04}
\scalebox{1}{
\begin{tabular}{ccccccccc}
  \hline
Simu. & Nat. REPNS Rate & New Nat. REPNS Rate & Mean & New Mean & Median & New Median & SD & New SD \\
  \hline
  1 & 0.9533 & 0.9467 & 0.9519 & 0.9469 & 0.9604 & 0.9545 & 0.0427 & 0.0291 \\
  2 & 0.9533 & 0.9467 & 0.9519 & 0.9466 & 0.9604 & 0.9545 & 0.0427 & 0.0344 \\
  3 & 0.9533 & 0.9467 & 0.9519 & 0.9467 & 0.9604 & 0.9544 & 0.0427 & 0.0301 \\
  4 & 0.9533 & 0.9467 & 0.9519 & 0.9466 & 0.9604 & 0.9545 & 0.0427 & 0.0302 \\
  5 & 0.9533 & 0.9467 & 0.9519 & 0.9468 & 0.9604 & 0.9545 & 0.0427 & 0.0291 \\
  6 & 0.9533 & 0.9467 & 0.9519 & 0.9466 & 0.9604 & 0.9543 & 0.0427 & 0.0304 \\
  7 & 0.9533 & 0.9467 & 0.9519 & 0.9467 & 0.9604 & 0.9544 & 0.0427 & 0.0298 \\
  8 & 0.9533 & 0.9466 & 0.9519 & 0.9467 & 0.9604 & 0.9545 & 0.0427 & 0.0342 \\
  9 & 0.9533 & 0.9467 & 0.9519 & 0.9467 & 0.9604 & 0.9546 & 0.0427 & 0.0296 \\
 10 & 0.9533 & 0.9466 & 0.9519 & 0.9468 & 0.9604 & 0.9546 & 0.0427 & 0.0292 \\
   \hline
\end{tabular}
}
\end{sidewaystable}
````

## Multiple rows in a table
````
\usepackage{multirow}

\begin{table}[htbp]
\label{no-lk-errors}
\begin{center}
\begin{tabular}{cc|cc}
& & \multicolumn{2}{c}{$S_2$}\\
& & present & absent\\
\hline
\multirow{2}{*}{$S_1$} & present & $x_{11}$ & $x_{12}$\\
& absent & $x_{21}$ & $-$
\end{tabular}
\end{center}
\end{table}
````


## Calligraphy fonts
````
$\mathcal{N}_0$
````

**Source:** https://tex.stackexchange.com/questions/84041/why-does-calm-n-give-m

## Type degree
````
$360^{\circ}$
````

**Source:** http://web.ift.uib.no/HP/LaTeX_degree.html


# R tips {-}

## Testing `character(0)` in `if` statement

```{r, eval=FALSE}
v <- character(0)
identical(v, character(0))  # returns TRUE

identical(v, "")           # returns FALSE
identical(v, numeric(0))   # returns also FALSE
```

**Source:** https://stackoverflow.com/questions/25640161/r-how-to-test-for-character0-in-if-statement

## Removing/keeping objects


**Q**: How to remove all objects but some from the workspace in R?

**A**:
```{r, eval=FALSE}
rm(list=ls()[! ls() %in% c("a","b")])
```


## Programming with dplyr
```{r, eval=FALSE}
library(dplyr)

cols <- c("am","gear")
grouping_vars <- rlang::syms(cols)

my_df <- 
  mtcars %>%
  group_by(!!!grouping_vars) %>%
  summarise(mean_cyl=mean(cyl))
```

## Add an extra column 
**Q**: How to add a column to a dataframe?

**A**:
```{r, eval=FALSE}
library(tibble) # for add_column
library(dplyr)

df <-
  data.frame(a = 1:5, b = 2:6, c = 3:7) %>%
  add_column(d = 4:8, .after = 2)
```

## Add an extra row
```{r, message=FALSE, warning=FALSE}
library(tibble) # for add_row

df_a <- data.frame(x = letters[8:10], y = 1:3)

(df_b <- add_row(df_a, x = "x", y = 10, .before = 1))

```


## De-duplicate a dataframe
```{r, message=FALSE, warning=FALSE}
library(dplyr)

a_df <- 
  data.frame(x = rep(letters[1:3], each = 2),
             y = rep(letters[2:4], each = 2),
             z = rep(1:2, each = 3))
(a_df)

b_df <- 
  a_df %>% 
  distinct()
(b_df)

c_df <- 
  a_df %>% 
  distinct(x, y, .keep_all = TRUE)
(c_df)
```


## Anti- and semi-join
```{r, message=FALSE, warning=FALSE}
library(dplyr)

a_df <- 
  data.frame(x = rep(letters[1:3], each = 2),
             y = rep(letters[2:4], each = 2),
             z = rep(1:2, each = 3))
(a_df)

c_df <- 
  a_df %>% 
  distinct(x, y, .keep_all = TRUE)
(c_df)

d_df <- semi_join(a_df, c_df)
(d_df)

e_df <- anti_join(a_df, c_df)
(e_df)

```



## Simple recoding

```{r, warning=FALSE, message=FALSE}
library(dplyr)

x <- letters[1:5]
y <- recode(x, a = "apple")
z <- recode(x, b = "banana")

c(y, z)

```

**Source**: https://dplyr.tidyverse.org/reference/recode.html

## Use `grepl()` to filter
```{r, warning=FALSE, message=FALSE}
library(dplyr)

df_1 <- 
  data.frame(x = c("x-18", sample(letters[1:3], 3), "a-18", "b-18"), y = rnorm(6))

# get the rows where x contains '-'
(df_2 <- filter(df_1, grepl("-", x)))

```





## Find number of NA's 
**Q**: How to check NA's in a dataframe?

**A**: `colSums(is.na(df))`

## Subset a dataframe by NA
**Q**: How to get the rows which have NA's in a dataframe?

**A**: `new_DF <- DF[rowSums(is.na(DF)) > 0, ]`

## Covert NA to 0
**Q**: How to convert NA to 0 in a dataframe?

**A**: `df[is.na(df)] <- 0`

## Convert blank to `NA`

```{r,  message=FALSE, warning=FALSE}
library(dplyr)

fake_df <-
  data.frame(x = c(1:2, "", 3, 4, ""),
             y = c("a", "b", "c", "", "", "d"),
             z = c(rep(TRUE, 4), "", ""))

(fake_df)

converted_fake_df <- 
  fake_df %>% 
  mutate_all(na_if, "")

(converted_fake_df)
```

**Source:** https://stackoverflow.com/questions/24172111/change-the-blank-cells-to-na



## Label y axis in ggplot2
**Q**: How to label y axis with percent/dollar/comma sign?

**A**: 
```{r, eval=FALSE}
p <- ggplot(df, aes(x, y)) + geom_point()
p + scale_y_continuous(labels = scales::percent)
p + scale_y_continuous(labels = scales::dollar)
p + scale_y_continuous(labels = scales::comma)
```

## Set plot margins in ggplot2
```{r, eval=FALSE}
library(ggplot2)

my_df <- data.frame(x = 1:10, y = 1:10)

ggplot(my_df, aes(x, y)) +
  geom_point() +
  theme(plot.margin = unit(c(1, 2, 4, 6), "cm")) # top, right, bottom and left
```

## Set plot margins in base R
```{r, eval=FALSE}
# sets the bottom, left, top and right margins respectively of 
# the plot region in number of lines of text
par(mar = c(10, 8, 4, 2))
plot(1:10, 1:10, xlab = "x", ylab = "y")
```

## Deal with long labels in ggplot2

**Source**: https://stackoverflow.com/questions/41568411/how-to-maintain-size-of-ggplot-with-long-labels

## Edit plotly object
```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(plotly)

df <- data.frame(x = c(1, 2, 3, 4), 
                 y = c(1, 5, 3, 5), 
                 group = c('A', 'A', 'B', 'B'))
g <- 
  ggplot(data=df, aes(x = x, y = y, colour = group)) + 
  geom_point(size = 4) +
  geom_hline(yintercept = 2) +
  scale_color_manual(values=c("red", "blue")) +
  theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"))

p <- plotly_build(g) %>% config(displayModeBar = FALSE)

# to have a look at object p
# str(p)

# modify the text from "~yintercept: 2" to "y-intercept = 2"
p$x$data[[3]]$text <- "y-intercept = 2"

p
```


## Zoom in on a barchart

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(gridExtra)

the_data <- data.frame(x = LETTERS[1:3], y = c(200, 300, 150))

# original
p1 <-
  ggplot(the_data, aes(x = x, weight = y)) +
  geom_bar(width = 0.5, fill = "blue") +
  labs(x = "")

# zoomed 
p2 <- 
  p1 + coord_cartesian(ylim = c(100, 300))

# horizontal bars; zoomed
p3 <-
  p1 + coord_flip(ylim = c(100, 300))

p4 <- ggplot()

# put tem together
grid.arrange(p1, p2, p3, p4, newpage = TRUE, 
             layout_matrix = matrix(1:4, byrow = TRUE, 2, 2))
```



## List loaded packages
**Q**: How to list the packages currently loaded into R's work space? 

**A**:  `(.packages())`

**Source**: http://theobligatescientist.blogspot.com/2010/09/r-tip-listing-loaded-packages.html


## Function names in a loaded package
**Q**: How to find function names in a loaded package?

**A**: Firstly load the package using `library()`, and then use `lsf.str(package:package_name)`. For example

```{r, eval=FALSE}
library(tidyr)
lsf.str("package:tidyr")
```

**Source**: https://stackoverflow.com/questions/30392542/is-there-a-command-in-r-to-view-all-the-functions-present-in-a-package

 


## Vectorize a matrix

```{r, eval=FALSE}
x <- matrix(1:10, 2, 5)
# vectorize it by columns
c(x) 
# or
as.vector(x)
# vectorize it by rows 
c(t(x))
```

**Source**: https://stackoverflow.com/questions/4570537/vectorizing-a-matrix


# Python tips {-}

## Viewing all defined variables
```{r, eval=FALSE}
dir() # gives the list of in scope variables:
globals() # gives a dictionary of global variables
locals() # gives a dictionary of local variables
```

**Source**: https://stackoverflow.com/questions/633127/viewing-all-defined-variables

## `type()`, `dir()`, and `help()`

The `type` function shows the type of an object; the `dir` function lists all the methods available for an object; `help` function gives explanation about a method.

**Example**:
```{r, eval=FALSE}
stuff = "Hello world"

type(stuff)
<class 'str'>

dir(stuff)


help(str.capitalize)

```

## Commenting out

* Single line comment: Ctrl + 1
* Multi-line comment: select the lines to be commented Ctrl + 4
* Unblock Multi-line comment: Ctrl + 5

Resource: https://stackoverflow.com/questions/36644144/shortcut-key-for-commenting-out-lines-of-python-code-in-spyder