--- 
title: "Useful Tips"
author: "Lingyun Zhang"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
link-citations: yes
links-as-notes: true
colorlinks: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preface {-}

This book contains useful Tex, R and Python tips.

# Tex tips {-}

## Commenting in WinEdt

In Winedt use `Ctrl+Shift+Alt+→` to comment a line and to uncomment a line use `Ctrl+Shift+Alt+←`.

**Source**: https://tex.stackexchange.com/questions/230997/which-windows-editors-have-comment-uncomment-keyboard-shortcuts


## Rotate a table

**Q**: How to rotate a table in latex?

**A**:
````
\usepackage{rotating}
\begin{sidewaystable}[h]
\centering
\caption{Mean, Median and Standard Deviation at AU level from Simulations, where
                  tolerance limit set at MB level and $p=0.04$}
\label{table04}
\scalebox{1}{
\begin{tabular}{ccccccccc}
  \hline
Simu. & Nat. REPNS Rate & New Nat. REPNS Rate & Mean & New Mean & Median & New Median & SD & New SD \\
  \hline
  1 & 0.9533 & 0.9467 & 0.9519 & 0.9469 & 0.9604 & 0.9545 & 0.0427 & 0.0291 \\
  2 & 0.9533 & 0.9467 & 0.9519 & 0.9466 & 0.9604 & 0.9545 & 0.0427 & 0.0344 \\
  3 & 0.9533 & 0.9467 & 0.9519 & 0.9467 & 0.9604 & 0.9544 & 0.0427 & 0.0301 \\
  4 & 0.9533 & 0.9467 & 0.9519 & 0.9466 & 0.9604 & 0.9545 & 0.0427 & 0.0302 \\
  5 & 0.9533 & 0.9467 & 0.9519 & 0.9468 & 0.9604 & 0.9545 & 0.0427 & 0.0291 \\
  6 & 0.9533 & 0.9467 & 0.9519 & 0.9466 & 0.9604 & 0.9543 & 0.0427 & 0.0304 \\
  7 & 0.9533 & 0.9467 & 0.9519 & 0.9467 & 0.9604 & 0.9544 & 0.0427 & 0.0298 \\
  8 & 0.9533 & 0.9466 & 0.9519 & 0.9467 & 0.9604 & 0.9545 & 0.0427 & 0.0342 \\
  9 & 0.9533 & 0.9467 & 0.9519 & 0.9467 & 0.9604 & 0.9546 & 0.0427 & 0.0296 \\
 10 & 0.9533 & 0.9466 & 0.9519 & 0.9468 & 0.9604 & 0.9546 & 0.0427 & 0.0292 \\
   \hline
\end{tabular}
}
\end{sidewaystable}
````

# R tips {-}

## Programming with dplyr
```{r, eval=FALSE}
library(dplyr)

cols <- c("am","gear")
grouping_vars <- rlang::syms(cols)

my_df <- 
  mtcars %>%
  group_by(!!!grouping_vars) %>%
  summarise(mean_cyl=mean(cyl))
```

## Add an extra column 
**Q**: How to add a column to a dataframe?

**A**:
```{r, eval=FALSE}
library(tibble) # for add_column
library(dplyr)

df <-
  data.frame(a = 1:5, b = 2:6, c = 3:7) %>%
  add_column(d = 4:8, .after = 2)
```

## Add an extra row
```{r, message=FALSE, warning=FALSE}
library(tibble) # for add_row

df_a <- data.frame(x = letters[8:10], y = 1:3)

(df_b <- add_row(df_a, x = "x", y = 10, .before = 1))

```



## Simple recoding

```{r, warning=FALSE, message=FALSE}
library(dplyr)

x <- letters[1:5]
y <- recode(x, a = "apple")
z <- recode(x, b = "banana")

c(y, z)

```

**Source**: https://dplyr.tidyverse.org/reference/recode.html

## Use `grepl()` to filter
```{r, warning=FALSE, message=FALSE}
library(dplyr)

df_1 <- 
  data.frame(x = c("x-18", sample(letters[1:3], 3), "a-18", "b-18"), y = rnorm(6))

# get the rows where x contains '-'
(df_2 <- filter(df_1, grepl("\\-", x)))

```





## Find number of NA's 
**Q**: How to check NA's in a dataframe?

**A**: `colSums(is.na(df))`

## Subset a dataframe by NA
**Q**: How to get the rows which have NA's in a dataframe?

**A**: `new_DF <- DF[rowSums(is.na(DF)) > 0, ]`

## Covert NA to 0
**Q**: How to convert NA to 0 in a dataframe?

**A**: `df[is.na(df)] <- 0`

## Label y axis in ggplot2
**Q**: How to label y axis with percent/dollar/comma sign?

**A**: 
```{r, eval=FALSE}
p <- ggplot(df, aes(x, y)) + geom_point()
p + scale_y_continuous(labels = scales::percent)
p + scale_y_continuous(labels = scales::dollar)
p + scale_y_continuous(labels = scales::comma)
```

## Set plot margins in ggplot2
```{r, eval=FALSE}
library(ggplot2)

my_df <- data.frame(x = 1:10, y = 1:10)

ggplot(my_df, aes(x, y)) +
  geom_point() +
  theme(plot.margin = unit(c(1, 2, 4, 6), "cm")) # top, right, bottom and left
```

## Set plot margins in base R
```{r, eval=FALSE}
# sets the bottom, left, top and right margins respectively of 
# the plot region in number of lines of text
par(mar = c(10, 8, 4, 2))
plot(1:10, 1:10, xlab = "x", ylab = "y")
```

## Deal with long labels in ggplot2

**Source**: https://stackoverflow.com/questions/41568411/how-to-maintain-size-of-ggplot-with-long-labels

## Edit plotly object
```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(plotly)

df <- data.frame(x = c(1, 2, 3, 4), 
                 y = c(1, 5, 3, 5), 
                 group = c('A', 'A', 'B', 'B'))
g <- 
  ggplot(data=df, aes(x = x, y = y, colour = group)) + 
  geom_point(size = 4) +
  geom_hline(yintercept = 2) +
  scale_color_manual(values=c("red", "blue")) +
  theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"))

p <- plotly_build(g) %>% config(displayModeBar = FALSE)

# to have a look at object p
# str(p)

# modify the text from "~yintercept: 2" to "y-intercept = 2"
p$x$data[[3]]$text <- "y-intercept = 2"

p
```


## List loaded packages
**Q**: How to list the packages currently loaded into R's work space? 

**A**:  `(.packages())`

**Source**: http://theobligatescientist.blogspot.com/2010/09/r-tip-listing-loaded-packages.html


## Vectorize a matrix

```{r, eval=FALSE}
x <- matrix(1:10, 2, 5)
# vectorize it by columns
c(x) 
# or
as.vector(x)
# vectorize it by rows 
c(t(x))
```

**Source**: https://stackoverflow.com/questions/4570537/vectorizing-a-matrix


# Python tips {-}

## Viewing all defined variables
```{r, eval=FALSE}
dir() # gives the list of in scope variables:
globals() # gives a dictionary of global variables
locals() # gives a dictionary of local variables
```

**Source**: https://stackoverflow.com/questions/633127/viewing-all-defined-variables

## `type()`, `dir()`, and `help()`

The `type` function shows the type of an object; the `dir` function lists all the methods available for an object; `help` function gives explanation about a method.

**Example**:
```{r, eval=FALSE}
stuff = "Hello world"

type(stuff)
<class 'str'>

dir(stuff)


help(str.capitalize)

```

